Augmentation of a Historical Harpsichord Keyboard Replica for Haptic-Enabled Interaction in Museum Exhibitions

### 2025-01-21 src/article.tex
Updates from Overleaf

\begin{figure}  
  \centering
  \includegraphics[width=\linewidth]{images/side_w-o_sensors.png} 
  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023} 
  \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
mechanism for an Italian style harpsichord of the XVI century.} 
  \label{fig:3key}
\end{figure}

@@

@@

Central philosophy of the Tagliavini Collection is that the instrument
should remain playable to the public. The condition of some instruments
is such that restoration would required a replacement of vast majority
of the instrument, would be too costly, would risk the instrument given
its fragility, or all of those in combination.

This is not an uncommon problem faced by musical instrument museums
worldwide. The central question the project tries to answer is ``how do
we conserve musical instruments in order to continue interacting with
them''

The approach taken was to recreate the interface of these instruments
not for acoustics purposes be purely for the mechanical response. As
exhibition for the general public in a musical instruments there were
aesthetic and functional constraints placed on the design of the
interface.

The design criteria was to create a keyboard interface that provides a
comparable physical response to a harpsichord.

@@

\begin{itemize}
\item
  no electronics should be visible.
\item
  it should be robust and reliable, easily maintained by museum staff
  without complex technical processes
\item
  it should not augment or fix the mechanics, but present them with
  their original limitation
\end{itemize}

Another constraint was that the design of the sensor system and the
model keyboard would have to carried out at a distance in separate
workshops. Time constraints meant the keyboard was being fabricated for
a sensor that did not yet exist.

Accommodations were made in the design to allow for as much flexibility
as possible. This also meant that designs for the electronics would have
to include a deal of flexibility in how they might be installed. Surplus
space was added into the design of the internal chamber of the model
\textbackslash FIGURE, but there was still a limit given the structural
and aesthetic requirements.

@@

The model was constructed using the materials of the Italian harpsichord
tradition:

@@

For the initial prototyping stage, modifying an existing harpsichord was
considered though ultimately discarded. An existing instrument would
have provided a test for scaling the electronics, but internal
measurements and layout would have been too different. A harpsichord that
was economic enough to be deconstructed would also have been of a more
modern mechanical design and not of the design that was being recreated
with the model. Prototyping stage would likely be invasive and damage or
heavy modification likely. As such it was not a good use of funds as a
modern instrument fit for destruction did not offer benefits that the
small model with easy internal access did not already provide.

@@

@@

Design of electronics carried out in tandem with construction of full
model. Since the Electronics were designed at a distanced to the final
full model, some flexibility was required in the design. Design decisions
that were not as optimal but would allow for change in wiring and
fitting should there be any issues during installation.

Electronics designed so that they would not require any equipment
outside what can commonly be found in a maker space at a university.

The equipment required including soldering suitable for through-hole
technology (THT) components and 3D printing.

Laser cutting and CNC routing facilities were used during the research
for the project but not required for the final implementation.

The idea behind this was to allow for the project to be reasonably
recreated without procurement of any additional specialist equipment.

NEMUS lab under construction during the first section of the project.
Facilities available at the maker spaces of \anon{Bologna (Alma Labor)} and
\anon{Edinburgh (uCreate)} used for what was deemed reasonable assumption in
terms of available equipment.

@@

The 49 QRE1113s are divided across 7 PCBs each containing

\subsubsection{QRE1113}\label{qre1113}

A number of sensors were tried before the final system was decided upon.
Sensors included, an 6-degrees-of-freedom inertial measurement unit,
reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
None of these sensors satisfied the sensor criteria described above.

They \emph{could} be used and individual trigger points set in
calibration, but this creates new problems.

An initial sticker design was used that contained only the gradient. It
was found that the jacks were sticking in their slots and the edges on the
stickers easily peeled around the printing The sticker was redesigned to
be the full length of the jack which avoided the problem of peeling. The
larger are also mean the stickers adhered better to the jack Some
trimming was still required of each sticker to avoid them adding extra
resistance to the jack's movement.

>>>>

\begin{figure}  
  \centering
  \includegraphics[width=\linewidth]{src/images/3-key-side-labelled.png} 
  \caption{3-key Model Harpsichord Mechanism by Graziano Bandini, Bologna 2023. 3-Key modelled used for initial prototyping. A: Wrestplank B: Key Lever C: Key cap D: soundboard E: Jack} 
  \Description{A 3-Key Model of a Harpsichord Mechanism created Graziano Bandini in Bologna. A side view of the model which is a cross-section of what is a common
mechanism for an Italian style harpsichord of the XVI century.} 
  \label{fig:3key}
\end{figure}

@@

The design criteria was to create a keyboard interface that provided a comparable tactile response to that of a harpsichord in the italian style.

Accommodations were made in the design to allow for as much flexibility as possible. 
This also meant that designs for the electronics would have to include a level of flexibility in how they might be installed. 
Surplus space was added into the design of the internal chamber of the model Figure \ref{fig:49-key-bottom}, but there was still a limit given the structural
and aesthetic requirements.

@@

% Central philosophy of the Tagliavini Collection is that the instrument
% should remain playable to the public. The condition of some instruments
% is such that restoration would required a replacement of vast majority
% of the instrument, would be too costly, would risk the instrument given
% its fragility, or all of those in combination.

% This is not an uncommon problem faced by musical instrument museums
% worldwide. The central question the project tries to answer is ``how do
% we conserve musical instruments in order to continue interacting with
% them''

% The approach taken was to recreate the interface of these instruments
% not for acoustics purposes be purely for the mechanical response. As
% exhibition for the general public in a musical instruments there were
% aesthetic and functional constraints placed on the design of the
% interface.

@@

There were constraints that were put in place by the Museum on the design of the exhibit.

\begin{itemize}
\item
  No electronics should be visible \footnote{with the exception of a pair of headphones for listening and a display to change parameters of the audio engine}
\item
  It should be robust and reliable, easily maintained by museum staff without complex technical processes
\item
  it should not augment or fix the mechanics, but present them with their original limitation
\end{itemize}

% Another constraint was that the design of the sensor system and the
% model keyboard would have to carried out at a distance in separate
% workshops. Time constraints meant the keyboard was being fabricated for
% a sensor that did not yet exist.

@@

The model was constructed using the materials of the Italian harpsichord
tradition. Parts are labelled in Figure \ref{fig:3key}.

@@

% A harpsichord that
% was economic enough to be deconstructed would also have been of a more
% modern mechanical design and not of the design that was being recreated
% with the model. Prototyping stage would likely be invasive and damage or
% heavy modification likely. As such it was not a good use of funds as a
% modern instrument fit for destruction did not offer benefits that the
% small model with easy internal access did not already provide.

@@

For the initial prototyping stage, modifying an existing harpsichord was considered though ultimately discarded. An existing instrument would
have provided a test for scaling the electronics, but internal measurements and layout would have been too different.

Using a similar approach to Timmermans' Haptic Key project, \cite{Timmermans2020} our first step was to fabricate or procure a model of a single working key. San Colombano was able to provide a model 3-key of a harpischord mechanism by Graziano Bandini (Figure \ref{fig:3key}).

@@

% Design of electronics carried out in tandem with construction of full
% model. Since the Electronics were designed at a distanced to the final
% full model, some flexibility was required in the design. Design decisions
% that were not as optimal but would allow for change in wiring and
% fitting should there be any issues during installation.

The sensor system was designed so that they would not require any equipment outside what can commonly be found in a maker space at a university for assembly.

% The equipment required including soldering suitable for through-hole technology (THT) components and 3D printing.

% Laser cutting and CNC routing facilities were used during the research
% for the project but not required for the final implementation.

This decisions was motivated by the project's goal to be open source and reasonably recreated without procurement of any additional specialist equipment.

% NEMUS lab under construction during the first section of the project.
% Facilities available at the maker spaces of \anon{Bologna (Alma Labor)} and
% \anon{Edinburgh (uCreate)} used for what was deemed reasonable assumption in
% terms of available equipment.

The electronics system can be divided into two designs, the "sensor board" (Figure \ref{fig:7-sensor-board}) for components for detecting jack displacement and controller board for collecting and processing sensor signals.

The 49 optical sensors are divided across 7 PCBs each containing

\begin{itemize}
\item
  x7 QRE1113 optical sensors
\item
  x7 100 $\Omega$ resistors
\item
  x7 10 k$\Omega$ resistors \footnote{a update on the design showed these could be reduced to a single resistor.}
\item
  x1 CD4051BE Multipexer
\end{itemize}

A number of sensors were tried before the final system was decided upon. 
Sensors tested for the design included, a 6-degrees-of-freedom inertial measurement unit, reed and hall sensors paired with magnets, and light-dependant resistors mounted to the jack rail.
None of the above satisfied the sensor criteria described and were thus discounted.

% They \emph{could} be used and individual trigger points set in
% calibration, but this creates new problems.

% An initial sticker design was used that contained only the gradient. It
% was found that the jacks were sticking in their slots and the edges on the
% stickers easily peeled around the printing The sticker was redesigned to
% be the full length of the jack which avoided the problem of peeling. The
% larger are also mean the stickers adhered better to the jack Some
% trimming was still required of each sticker to avoid them adding extra
% resistance to the jack's movement.

@@
